---
title: "BreakoutICU"
author: "Deniz Avci"
date: "2025-06-25"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(rpart)
library(rpart.plot)
library(rattle)
icu_data = read.csv("ICU.csv")
icu_train = icu_data[1:150,]
icu_test = icu_data[151:200,]
attach(icu_train)

my_tree = rpart(formula = Survive ~ Age + AgeGroup + Sex + Infection + SysBP + 
                  Pulse + Emergency, data = icu_train, method = "class", cp = 0.015)

rpart.plot(my_tree)
```

Those with systolic bloodpressure below 91 have a lower survival rate (33%) than
those with higher than 91 (82%). Additionaly, patients who didnt consulted the 
emergency room prior to thier ICU visit had a 96% survival rate, whereas those who
did, ex. for a gunshot wound, car crash, or sudden event, had a 75% rate.

```{r}
my_prediction = predict(my_tree, icu_test, type = "class")
table(my_prediction, icu_test$Survive)
mean(my_prediction == icu_test$Survive)
```
Here we can see that the 0.82 represents the accuracy of our model: 82% correct.

```{r}
library(randomForest)
my_forest <- randomForest(factor(Survive) ~ Age + AgeGroup + Sex + Infection + SysBP
                          + Pulse + Emergency, data= icu_train, 
                          importance = TRUE, ntree = 1000)

varImpPlot(my_forest)
```
By focusing on the MeanDecreaseAccuracy plot, the Infection predictor stands out
as a variable of least importance, meaning that its removal from the model could
improve efficiency if it truly has no impact on our accuracy result. We test this
below.

```{r}
my_tree2 = rpart(formula = Survive ~ Age + AgeGroup + Sex + SysBP + 
                  Pulse + Emergency, data = icu_train, method = "class", cp = 0.015)

rpart.plot(my_tree2)
```

```{r}
my_prediction2 = predict(my_tree2, icu_test, type = "class")
table(my_prediction2, icu_test$Survive)
mean(my_prediction2 == icu_test$Survive)
```
Because there was no change in %accuracy between the first run and this run, the 
removal of predictor: "Infection" does not impact the data substantially, therefore
its removal increases the efficiency of the program.